---
title: "Use Python with R with reticulate :: Cheat Sheet"
format: 
  html:
    toc: true
    highlight-style: a11y-dark
editor: visual
---

The **reticulate** package lets you use Python and R together seamlessly in R code, in R Markdown documents, and in the RStudio IDE.

```{r}
library(reticulate)
```

## Python in R Markdown

(Optional) Build Python env to use.

Add `knitr::knit_engires$set(python = reticulate::eng_python)` to the setup chunk to set up the reticulate Python engine (not required for knitr \>= 1.18).

````         
```{{r}}
#| label: setup
#| include: false

library(reticulate) 
virtualenv_create("fmri-proj") 
py_install("seaborn", envname = "fmri-proj") 
use_virtualenv("fmri-proj")`
```
````

Begin Python chunks with ```` ```{python} ````.
Chunk options like `echo`, `include`, etc. all work as expected.

````         
```{{python}} 
#| echo: false

import seaborn as sns 
fmri = sns.load_dataset("fmri")`
```
````

Use the `py` object the access objects created in Python chunks from R chunks.

``` {{r}}
f1 <- subset(py$fmri, region = "parietal")
```

Python chunks all execute within a **single** Python session so you have access to all objects created in previous chunks.

Use the `r` object to access objects created in R chunks from Python chunks.

``` {{python}}
import matplotlib as mpl
sns.lmplot("timepoint", "signal", data=r.f1)
mpl.pyplot.show()
```

## Python in R

`python.r` Example:

```{r}
#| eval: false

library(reticulate)
py_install("seaborn")
use_virtualenv("r-reticulate")

sns <- import("seaborn")

fmri <- sns$load_dataset("fmri")
dim(fmri)

# creates tips
source_python("python.py")
dim(tips)

# creates tips in main
py_run_file("python.py")
dim(py$tips)

py_run_string("prints(tips.shape)")
```

Call Python from R code in three ways:

### Import Python Modules

Use `import()` to import any Python module.
Access the attributes of a module with `$`.

```{r}
#| eval: false

library(reticulate)
py_install("seaborn")
use_virtualenv("r-reticulate")

sns <- import("seaborn")

fmri <- sns$load_dataset("fmri")
dim(fmri)
```

-   `import(module, as = NULL, convert = TRUE, delay_load = FALSE)`: Import a Python module.
    If `convert = TRUE`, Python objects are converted to their equivalent R types.
    Also `import_from_path()`.

    ```{r}
    #| eval: false

    import("pandas")
    ```

-   `import_main(convert = TRUE)`: Import the main module, where Python executes code by default.

    ```{r}
    #| eval: false

    import_main()
    ```

-   `import_builtings(convert = TRUE)`: Import Python's built-in functions.

    ```{r}
    #| eval: false

    import_builtins()
    ```

### Source Python Files

Use `source_python()` to source a Python script and make the Python functions and objects it creates available in the calling R environment.

```{r}
#| eval: false

source_python("python.py")
dim(py$tips)
```

-   `source_python(file, envir = parent.frame(), convert = TRUE)`: Run a Python script, assigning objects to a specified R environment.

    ```{r}
    #| eval: false

    source_python("file.py")
    ```

### Run Python Code

Execute Python code into the `main` Python modules with `py_run_file()` or `py_run_string()`.

```{r}
#| eval: false

py_run_file("python.py")
dim(py$tips)

py_run_string("print(tips.shape)")
```

-   `py_run_string(code, local = FALSE, convert = TRUE)`: Run Python code (passed as a string) in the main module.

    ```{r}
    #| eval: false

    py_run_string("x = 10")
    py$x
    ```

-   `py_run_file(file, local = FALSE, convert = TRUE)`: Run Python file in the main module.

    ```{r}
    #| eval: false

    py_run_file("script.py")
    ```

-   `py_eval(code, convert = TRUE)`: Run a Python expression, return the result.
    Also `py_call()`.

    ```{r}
    #| eval: false

    py_eval("1 + 1")
    ```

Access the results, and anything else in Python's `main` module, with `py`.

-   `py`: An R object that contains the Python main module and the results stored there.

    ```{r}
    #| eval: false

    py$x
    ```

## Object Conversion

**Tip: To index Python objects begin at 0, use integers, e.g. `OL`**

Reticulate provides **automatic** built-in conversion between Python and R for many Python types.

| R                      | Python            |
|------------------------|-------------------|
| Single-element vector  | Scalar            |
| Multi-element vector   | List              |
| List of multiple types | Tuple             |
| Named List             | Dict              |
| Matrix/Array           | NumPy ndarray     |
| Data Frame             | Pandas DataFrame  |
| Function               | Python function   |
| NULL, TRUE, FALSE      | None, True, False |

: Table of data types in R and their Python equivalents.

Or, if you like, you can convert manually with

-   `py_to_r(x)`: Convert a Python object to an R object.
    Also `r_to_py()`.

    ```{r}
    #| eval: false

    py_to_r(x)
    ```

-   `tuple(..., convert = FALSE)`: Create a Python tuple.

    ```{r}
    #| eval: false

    tuple("a", "b", "c")
    ```

-   `dict(..., convert = FALSE)`: Create a Python dictionary object.
    Also `py_dict()` to make a dictionary that uses Python objects as keys.

    ```{r}
    #| eval: false

    dict(foo = "bar", index = 42L)
    ```

-   `np_array(data, dtype = NULL, order = "C")`: Create NumPy arrays.

    ```{r}
    #| eval: false

    np_array(c(1:8), dtype = "float16")
    ```

-   `array_reshape(x, dim, order = c("C", "F"))`: Reshape a Python array.

    ```{r}
    #| eval: false

    x <- 1:4
    array_reshape(x, c(2,2))
    ```

-   `py_func(f)`: Wrap an R function in a Python function with the same signature.

    ```{r}
    #| eval: false

    py_fuc(xor)
    ```

-   `py_main_thread_func(f)`: Create a function that will always be called on the main thread.

-   `iterate(it, f = base::identity, simplify = TRUE)`: Apply an R function to each value of a Python iterator or return the values as an R vector, draining the iterator as you go.
    Also `iter_next()` and `as_iterator()`.

    ```{r}
    #| eval: false

    iterate(iter, print)
    ```

-   `py_interator(fn, completed = NULL)`: Create a Python iterator from an R function.

    ```{r}
    #| eval: false

    seq_gen <- function(x) {
      n <- x;
      function() {
        n <<- n + 1;
        n
      }
    }
    py_iterator(seq_gen(9))
    ```

## Helpers

-   `py_capture_output(expr, type = c("stdout", "stderr"))`: Capture and return Python output.
    Also `py_supress_warnings()`.

    ```{r}
    #| eval: false

    py_capture_output("x")
    ```

-   `py_get_attr(x, name, silent = FALSE)`: Get an attribute of a Python object.
    Also `py_set_attr()`, `py_has_attr()`, and `py_list_attributes()`.

    ```{r}
    #| eval: false

    py_get_attr(x)
    ```

-   `py_help(object)`: Open the documentation page for a Python object.

    ```{r}
    #| eval: false

    py_help(sns)
    ```

-   `py_last_error()`: Get the last Python error encountered.
    Also `py_clear_last_error()` to clear the last error.

    ```{r}
    #| eval: false

    py_last_error()
    ```

-   `py_save_object(object, filename, pickle = "pickle", ...)`: Save and load Python objects with pickle.
    Also `py_load_object()`.

    ```{r}
    #| eval: false

    py_save_objects(x, "x.pickle")
    ```

-   `with(data, expr, as = NULL, ...)`: Evaluate an expression within a Python context manager.

    ```{r}
    #| eval: false

    py <- import_builtins()
    with(py$open("output.txt", "w") %as% file,
         {file$write("Hello, there!")})
    ```

<!-- Page 2 -->

## Python in the IDE

Requires reticulate plus RStudio v1.2+.
Some features require v1.4+.
Screenshot Here

Expand to read about the reticulate features in the RStudio IDE.

#### Reticulate features in the RStudio IDE

-   Syntax highlighting for Python scripts and chunks.
-   Tab completion for Python functions and objects (and Python modules imported in R scripts).
-   Source Python scripts.
-   Execute Python code line by line with `Cmd + Enter` (`Ctrl + Enter`).
-   View Python objects in the Environment Pane.
-   View Python objects in the Data Viewer.
-   A Python REPL opens in the console when you run Python code with a keyboard shortcut. Type `exit` to close.
-   `matplotlib` plots display in plots pane.
-   Press `F1` over a Python symbol to display the help topic for that symbol.

## Python REPL

RStudio IDE Window:

A REPL (Read, Eval, Print Loop) is a command line where you can run Python code and view the results.

1.  Open in the console with `repl_python()`, or by running code in a Python script with `Cmd + Enter` (`Ctrl + Enter`).

    -   `repl_python(module = NULL, quiet = getOption("reticulate.repl.quiet", default = FALSE), input = NULL)`: Launch a Python REPL. Run `exit` to close.

        ```{r}
        #| eval: false

        repl_python()
        ```

2.  Type commands at `>>>` prompt.

3.  Press `Enter` to run code.

4.  Type `exit` to close and return to R console.

    ```         
    > reticulate::repl_python()
    Python 3.6.13 (/Users/rstudiointern/Library/r-miniconda/envs/r-reticulate/bin/python)
    Reticulate 1.20 REPL -- A Python interpreter in R.
    >>> inport seaborn as sns
    >>> tips = sns.load_dataset("tips")
    >>> tips.shape
    (244, 7)
    >>> exit
    >
    ```

## Configure Python

Reticulate binds to a local instance of Python when you first call `import()` directly or implicitly from an R session.
To control the process, find or build your desired Python instance.
Then suggest your instance to reticulate.
**Restart R to unbind.**

### Find Python

-   `install_python(version, list = FALSE, force = FALSE)`: Download and install Python.

    ```{r}
    #| eval: false

    install_python("3.6.13")
    ```

-   `py_available(initialize = FALSE)`: Check if Python is available on your system.
    Also `py_module_available()` and `py_numpy_module()`.

    ```{r}
    #| eval: false

    py_available()
    ```

-   `py_discover_config()`: Return all detected version of Python.
    Use `py_config()` to check which version has been loaded.

    ```{r}
    #| eval: false

    py_config()
    ```

-   `virtualenv_list()`: List all available virtualenvs.
    Also `virtualenv_root()`.

    ```{r}
    #| eval: false

    virtualenv_list()
    ```

-   `conda_list(conda = "auto")`: List all available conda envs.
    Also `conda_binary()` and `conda_version()`.

    ```{r}
    #| eval: false

    conda_list()
    ```

### Create a Python env

-   `virtual_env(envname = NULL, ...)`: Create a new virtual environment.

    ```{r}
    #| eval: false

    virtualenv_create("r-pandas")
    ```

-   `conda_create(envname = NULL, ...)`: Create a new conda environment.

    ```{r}
    #| eval: false

    conda_create("r-pandas", packages = "pandas")
    ```

### Install Packages

Install Python packages with R (below) or the shell:

`pip install SciPy`

`conda install SciPy`

-   `py_install(packages, envname, ...)`: Installs Python packages into a Python env.

    ```{r}
    #| eval: false

    py_install("pandas")
    ```

-   `virtualenv_install(envname, packages, ...)`: Install a package within a virtualenv.
    Also `virtualenv_remove()`.

    ```{r}
    #| eval: false

    virtualenv_install("r-pandas", packages = "pandas")
    ```

-   `conda_installs(envname, packages, ...)`: Install a package within a conda env.
    Also `conda_remove()`.

    ```{r}
    #| eval: false

    conda_install("r-pandas", packages = "plotly")
    ```

### Suggest an env to use

Set a default Python interpreter in the RStudio IDE Global or Project Options.
Go to **Tools \> Global Options ... \> Python** for Global Options.
Within a project, go to **Tools \> Project Options... \> Python**.

Otherwise, to choose an instance of Python to bind to, reticulate scans the instances on your computer in the following order, **stopping at the first instance that contains the module called by `import()`**.

1.  The instance referenced by the environment variable `RETICULATE_PYTHON` (if specified).
    **Tip: set in .Renviron file.**

    -   `Sys.setenv(RETICULATE_PYTHON = PATH)`: Set default Python binary.
        Persists across sessions!
        Undo with `Sys.unsetenv()`.

        ```{r}
        #| eval: false

        Sys.setenv(RETICULATE_PYTHON = "/usr/local/bin/python")
        ```

2.  The instances referenced by `use_` functions if called before `import()`.
    Will fail silently if called after impoty unless `required = TRUE`.

    -   `use_python(python, required = FALSE)`: Suggest a Python binary to use by path.

        ```{r}
        #| eval: false

        use_python("usr/local/bin/python")
        ```

    -   `use_virtualenv(virtualenv = NULL, required = FALSE)`: Suggest a Python virtualenv

        ```{r}
        #| eval: false

        use_virtualenv("~/myenv")
        ```

    -   `use_condaenv(condaenv = NULL, conda = "auto", required = FALSE)`: Suggest a conda env to use.

        ```{r}
        #| eval: false

        use_condaenv(condaenv = "r-nlp", conda = "/opt/anaconda3/bin/conda")
        ```

3.  Within virtualenvs and conda envs that carry the same name as the imported module.
    e.g. `~/anaconda/envs/nltk` for `import("nltk")`

4.  At the location of the Python binary discovered on the system PATH (i.e. `Sys.which("python")`)

5.  At customary locations for Python, e.g. `/usr/local/bin/python`, `/opt/local/bin/python…`

------------------------------------------------------------------------

CC BY SA Posit Software, PBC • [info\@posit.co](mailto:info@posit.co) • [posit.co](https://posit.co)

Learn more at [rstudio.github.io/reticulate](https://rstudio.github.io/reticulate/).

Updated: `r format(Sys.Date(), "%Y-%m")`.

```{r}
#| output: true

packageVersion("reticulate")
```

------------------------------------------------------------------------
